---
title: "Modelling"
output: html_document
date: "2023-09-22"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ATE analysis

Load libraries

```{r}
library(riskRegression)
library(dplyr)
library(survival)
library(Publish)
library(rms)
library(rsample)
library(lubridate)
library(xlsx)
library(data.table)
```

#### Load data frame

```{r}
df <- read.csv(".....csv")
```

#### Set covariates as factors 

```{r}
df$children_n.factor <- as.factor(df$children_n.factor)
df$comorbidities_factor <- as.factor(df$comorbidities_factor)
df$immigration_status <- as.factor(df$immigration_status)
df$bereaved01 <- as.factor(df$bereaved01)
df$sex <- as.factor(df$sex)
df$age_bereavement.factor <- as.factor(df$age_bereavement.factor)
df$affluence.factor <- as.factor(df$affluence.factor)


df$surv_cancer <- as.factor(df$surv_cancer)
df$surv_cardiovascular <- as.factor(df$surv_cardiovascular)
df$surv_neuro <- as.factor(df$surv_neuro)
df$surv_dm <- as.factor(df$surv_dm)
df$surv_digestive <- as.factor(df$surv_digestive)
df$surv_psychiatric <- as.factor(df$surv_psychiatric)
df$surv_resp <- as.factor(df$surv_resp)
```

#### Check for na

```{r}
sum(is.na(df$children_n.factor))
sum(is.na(df$comorbidities_factor))
sum(is.na(df$immigration_status))
sum(is.na(df$average_homecare))
sum(is.na(df$average_primarycare))
sum(is.na(df$average_residential))
sum(is.na(df$average_outpatient))
sum(is.na(df$average_inpatient))
sum(is.na(df$sex))
sum(is.na(df$bereaved01))
```

# Average bereavement effect (ATE)

In this section we model the ATE for the 7 causes of death: cancer, cardiovascular disease, dementia or Parkinsons, diabetes, digestive diseases, psychiatric diseases or suiside and respiratory diseases.

#### Distribution of covariates

For this specific analysis we used a categorized version of the expenditures data.

```{r}
df <- df %>%
  mutate(average_full_factor = case_when(average_full<0.01 ~ 1, #none
                                        average_full >=0.01 & average_full < 1 ~ 2, #low
                                        average_full >=1 & average_full <5 ~ 3, #some
                                        average_full>=5 ~ 4)) #high

df$average_full_factor <- as.factor(df$average_full_factor)

```

Look at the distribution of the covariates

```{r}
summary(utable(bereaved01 ~ sex + age_bereavement.factor + immigration_status +
                            comorbidities_factor + affluence.factor + 
                            children_n.factor +average_full_factor, 
                            data = df), 
                            show.totals = T)


```

## ATE: Sex stratified analysis

The data is split into seperate daat frames for males and females

```{r}
#New dataframe with only males
df_males <- df |> filter(sex == "1")

#New dataframe with only females
df_females <- df |> filter(sex == "2")

df_males$bereaved01 <- as.factor(df_males$bereaved01)
df_females$bereaved01 <- as.factor(df_females$bereaved01)

```

#### ABE, Cancer

-   The *ate* function is the average treatment effect. The first term is the formula from the CSC model.

```{r}
############ Males ###########
#cause-specific cox model (CSC)
csc_bereavement_cancer_males <- CSC(Hist(surv_time, surv_cancer) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                    data = df_males, 
                                    cause = "1")

#average treatment effect (ATE)
arv_bereavement_cancer_males <- ate(event = csc_bereavement_cancer_males, 
                                    treatment = "bereaved01", 
                                    data = df_males, 
                                    cause = "1", 
                                    times = c(365*3), 
                                    estimator = "GFORMULA", 
                                    se = TRUE, 
                                    B = 1000, 
                                    mc.cores = 15)


#show results
ABE_cancer_males <- print(setkeyv(as.data.table(arv_bereavement_cancer_males, type = c("meanRisk", "diffRisk")),"time"))


write.xlsx(ABE_cancer_males, "....xlsx")


################ Females ###################
#csc
csc_bereavement_cancer_females <- CSC(Hist(surv_time, surv_cancer) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                      data = df_females, 
                                      cause = "1")

#ate
arv_bereavement_cancer_females <- ate(event = csc_bereavement_cancer_females, 
                                      treatment = "bereaved01", 
                                      data = df_females, 
                                      cause = "1", 
                                      times = c(365*3), 
                                      estimator = "GFORMULA",  
                                      se = TRUE, 
                                      B = 1000, 
                                      mc.cores = 15)


#show results
ABE_cancer_females <- print(setkeyv(as.data.table(arv_bereavement_cancer_females, type = c("meanRisk", "diffRisk")),"time"))


write.xlsx(ABE_cancer_females, ".....xlsx")

```

#### ABE, cardio and vascular disease

```{r}
############ Males ###########

#csc
csc_bereavement_cardiovascular_males <- CSC(Hist(surv_time, surv_cardiovascular) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, data = df_males, cause = "1")

#ate
arv_bereavement_cardiovascular_males <- ate(event = csc_bereavement_cardiovascular_males, 
                                            treatment = "bereaved01", 
                                            data = df_males, 
                                            cause = "1", 
                                            times = c(365*3), 
                                            estimator = "GFORMULA",  
                                            se = TRUE, 
                                            B = 1000, 
                                            mc.cores = 15)


#show results
ABE_cardiovascular_males <- print(setkeyv(as.data.table(arv_bereavement_cardiovascular_males, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_cardiovascular_males, ".....xlsx")




################ Females ###################

#csc
csc_bereavement_cardiovascular_females <- CSC(Hist(surv_time, surv_cardiovascular) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                              data = df_females, 
                                              cause = "1")

#ate
arv_bereavement_cardiovascular_females <- ate(event = csc_bereavement_cardiovascular_females, 
                                              treatment = "bereaved01", 
                                              data = df_females, 
                                              cause = "1", 
                                              times = c(365*3), 
                                              estimator = "GFORMULA", 
                                              se = TRUE, 
                                              B = 1000, 
                                              mc.cores = 15)


#show results
ABE_cardiovascular_females <- print(setkeyv(as.data.table(arv_bereavement_cardiovascular_females, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_cardiovascular_females, "......xlsx")

```

#### ABE, Dementia and Parkinsons

```{r}

############ Males ###########

#csc
csc_bereavement_neuro_males <- CSC(Hist(surv_time, surv_neuro) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, data = df_males, cause = "1")

#ate
arv_bereavement_neuro_males <- ate(event = csc_bereavement_neuro_males, 
                                   treatment = "bereaved01", 
                                   data = df_males, 
                                   cause = "1", 
                                   times = c(365*3), 
                                   estimator = "GFORMULA", 
                                   se = TRUE, 
                                   B = 1000, 
                                   mc.cores = 15)


#show results
ABE_neuro_males <- print(setkeyv(as.data.table(arv_bereavement_neuro_males, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_neuro_males, ".....xlsx")


################ Females ###################
#csc
csc_bereavement_neuro_females <- CSC(Hist(surv_time, surv_neuro) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                     data = df_females, 
                                     cause = "1")

#ate
arv_bereavement_neuro_females <- ate(event = csc_bereavement_neuro_females, 
                                     treatment = "bereaved01", 
                                     data = df_females, 
                                     cause = "1", 
                                     times = c(365*3), 
                                     estimator = "GFORMULA", 
                                     se = TRUE, 
                                     B = 1000, 
                                     mc.cores = 15)


#show results
ABE_neuro_females <- print(setkeyv(as.data.table(arv_bereavement_neuro_females, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_neuro_females, "......xlsx")


```

#### ABE, Diabetes

```{r}
############ Males ###########

#csc
csc_bereavement_dm_males <- CSC(Hist(surv_time, surv_dm) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                data = df_males, 
                                cause = "1")

#ate
arv_bereavement_dm_males <- ate(event = csc_bereavement_dm_males, 
                                treatment = "bereaved01", 
                                data = df_males, 
                                cause = "1", 
                                times = c(365*3), 
                                estimator = "GFORMULA", 
                                se = TRUE, 
                                B = 1000, 
                                mc.cores = 15)



#show results
ABE_dm_males <- print(setkeyv(as.data.table(arv_bereavement_dm_males, type = c("meanRisk", "diffRisk")),"time"))


write.xlsx(ABE_dm_males, ".....xlsx")


################ Females ###################


#csc
csc_bereavement_dm_females <- CSC(Hist(surv_time, surv_dm) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                  data = df_females, 
                                  cause = "1")

#ate
arv_bereavement_dm_females <- ate(event = csc_bereavement_dm_females, 
                                  treatment = "bereaved01", 
                                  data = df_females, 
                                  cause = "1", 
                                  times = c(365*3), 
                                  estimator = "GFORMULA", 
                                  se = TRUE, 
                                  B = 1000, 
                                  mc.cores = 15)


#show results
ABE_dm_females <- print(setkeyv(as.data.table(arv_bereavement_dm_females, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_dm_females, ".....xlsx")
```

#### ABE, Diseases related to the digestive system

```{r}
############ Males ###########

#csc
csc_bereavement_digestive_males <- CSC(Hist(surv_time, surv_digestive) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                       data = df_males, 
                                       cause = "1")

#ate
arv_bereavement_digestive_males <- ate(event = csc_bereavement_digestive_males, 
                                       treatment = "bereaved01", 
                                       data = df_males, 
                                       cause = "1", 
                                       times = c(365*3), 
                                       estimator = "GFORMULA", 
                                       se = TRUE, 
                                       B = 1000, 
                                       mc.cores = 15)



#show results
ABE_digestive_males <- print(setkeyv(as.data.table(arv_bereavement_digestive_males, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_digestive_males, ".....xlsx")

################ Females ###################

#csc
csc_bereavement_digestive_females <- CSC(Hist(surv_time, surv_digestive) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                         data = df_females, 
                                         cause = "1")

#ate
arv_bereavement_digestive_females <- ate(event = csc_bereavement_digestive_females, 
                                         treatment = "bereaved01", 
                                         data = df_females, 
                                         cause = "1", 
                                         times = c(365*3),
                                         estimator = "GFORMULA", 
                                         se = TRUE, 
                                         B = 1000, 
                                         mc.cores = 15)


#show results
ABE_digestive_females <- print(setkeyv(as.data.table(arv_bereavement_digestive_females, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_digestive_females, "....xlsx")

```

#### ABE, Psychiatric diseases and suicide

```{r}
############ Males ###########
#csc
csc_bereavement_psychiatric_males <- CSC(Hist(surv_time, surv_psychiatric) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                         data = df_males, 
                                         cause = "1")

#ate
arv_bereavement_psychiatric_males <- ate(event = csc_bereavement_psychiatric_males, 
                                         treatment = "bereaved01", 
                                         data = df_males, 
                                         cause = "1", 
                                         times = c(365*3), 
                                         estimator = "GFORMULA", 
                                         se = TRUE, 
                                         B = 1000, 
                                         mc.cores = 15)



#show results
ABE_psychiatric_males <- print(setkeyv(as.data.table(arv_bereavement_psychiatric_males, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_psychiatric_males, "....xlsx")


################ Females ###################

#csc
csc_bereavement_psychiatric_females <- CSC(Hist(surv_time, surv_psychiatric) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                           data = df_females, 
                                           cause = "1")

#ate
arv_bereavement_psychiatric_females <- ate(event = csc_bereavement_psychiatric_females, 
                                           treatment = "bereaved01", 
                                           data = df_females, 
                                           cause = "1", 
                                           times = c(365*3), 
                                           estimator = "GFORMULA", 
                                           se = TRUE, 
                                           B = 1000, 
                                           mc.cores = 15)


#show results
ABE_psychiatric_females <- print(setkeyv(as.data.table(arv_bereavement_psychiatric_females, type = c("meanRisk", "diffRisk")),"time"))


write.xlsx(ABE_psychiatric_females, ".....xlsx")
```

#### ABE, Respiratory diseases

```{r}
############ Males ###########

#csc
csc_bereavement_resp_males <- CSC(Hist(surv_time, surv_resp) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                  data = df_males, 
                                  cause = "1")

#ate
arv_bereavement_resp_males <- ate(event = csc_bereavement_resp_males, 
                                  treatment = "bereaved01", 
                                  data = df_males, 
                                  cause = "1", 
                                  times = c(365*3), 
                                  estimator = "GFORMULA", 
                                  se = TRUE, 
                                  B = 1000, 
                                  mc.cores = 15)



#show results
ABE_resp_males <- print(setkeyv(as.data.table(arv_bereavement_resp_males, type = c("meanRisk", "diffRisk")),"time"))


write.xlsx(ABE_resp_males, ".....xlsx")

################ Females ###################

#csc
csc_bereavement_resp_females <- CSC(Hist(surv_time, surv_resp) ~ bereaved01 +  age_bereavement.factor + children_n.factor + average_full_factor  + affluence.factor + children_n.factor + comorbidities_factor + immigration_status, 
                                    data = df_females, 
                                    cause = "1")

#ate
arv_bereavement_resp_females <- ate(event = csc_bereavement_resp_females, 
                                    treatment = "bereaved01", 
                                    data = df_females, 
                                    cause = "1", 
                                    times = c(365*3), 
                                    estimator = "GFORMULA", 
                                    se = TRUE, 
                                    B = 1000, 
                                    mc.cores = 15)


#show results
ABE_resp_females <- print(setkeyv(as.data.table(arv_bereavement_resp_females, type = c("meanRisk", "diffRisk")),"time"))

write.xlsx(ABE_resp_females, "......xlsx")
```
